\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2023-02-22 Latex/paper.tex
# last version

In order to have a Markovian representation of the HJM, we will assume that the volatility is separable, i.e.
\begin{equation}
\sigma(t,T)= h(t)g(T).
\end{equation}
with $g$ a positive time-dependent function and $h$ a non-negative process. In addition
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Throught this paper, we assume the following hypotheses on $\eta(t,x,y)$.
\begin{hypothesis}\label{boundedness_volatility} 
The process $\eta_t$ is global Lipschitz and differentiable a.s. In addition, we will suppose that
\begin{align*}
a1 \leq \eta(t,x,y) \leq a2 \quad \forall (t,x,t) \in \mathbb{R}^{+} \times \mathbb{R} \times \mathbb{R}^{+}. \\
|\eta(t,x_2,y_2) - \eta(t,x_2,y_2)| \leq C_{x,y}\left(|x_2 - x_1| + |y_2 - y_1| \right) \quad \forall (t,x,t) \in \mathbb{R}^{+} \times \mathbb{R} \times \mathbb{R}^{+} 
\end{align*}
\end{hypothesis}
\begin{hypothesis}\label{boundedness_reversion} 
The next integrals are finite
\begin{align*}
I_{1,k}(0,t) &= \int_{0}^{t} \exp\left(- 2 \int_{s}^{t} k(s^{\prime})ds^{\prime}\right)  ds < \infty \\
I_{2,k}(0,t) &= \int_{0}^{t} \exp\left(- \int_{s}^{t} k(s^{\prime})ds^{\prime}\right) G(s,t) ds < \infty
\end{align*}
\end{hypothesis}
\begin{remark}
Notice that the above hypotheses have been chosen for the sake of simplicity, but they can be replaced by adequate integrability conditions.
\end{remark}
We must note, that under the hypothesis (\ref{boundedness_volatility}) $\partial_x \eta(t,x,y)$ and $\partial_y \eta(t,x,y)$ are bounded. This version of the HJM is also known as the Cheyette model, \cite{Cheyette}. It is easy to show (see \cite{Andreasen01} that
\begin{equation}
r_{ois}(t)=f_{ois}(t,t)= f_{ois}(0,t) + x_t,
\end{equation}  
where
\begin{align}\label{short_rate_cheyette}
dx_t &= (-k_t x_t + y_t)dt + \eta_t dW_t^{\mathbb{Q}} \nonumber \\
dy_t &= (\eta^{2}_t - 2 k_t y_t) dt ,
\end{align} 
with 
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Now if we do a little bit of algebra and we use Itô formula we can show the next representation formula for $P_{ois}(t,T)$
\begin{equation}\label{bond_ois}
P_{ois}(t,T) = \frac{P_{ois}(0,T)}{P_{ois}(0,t)} \exp\left(-G(t,T)x_t - \frac{1}{2} G^{2}(t,T)y_t \right)
\end{equation}
where $G(t,T) = \int_{t}^{T} \exp\left(-\int_{t}^{u} k_s ds \right) du$. We must observe, that under the representation (\ref{estimation_forward_rate_curve}) we have that
\begin{equation}\label{bond_forward}
P_{E}(t,T)=H(t,T)P_{ois}(t,T)
\end{equation}
where $H(t,T)=\exp\left(-\int_{t}^{T}s(t,u) du \right)$.

@@

\section{Order of the approximation $\bar{y}_t$}
Under the assumption of $\eta(t,\cdot,\cdot)$ is global Lipschitz and with first derive continuous, we have that
\begin{align*}
|\eta^{2}(t,x_t,y_t) - \eta^{2}(t,x_0,y_0)|&=|\eta(t,x_t,y_t) - \eta(t,x_0,y_0)||\eta(t,x_t,y_t) + \eta(t,x_0,y_0)| \nonumber \\ 
&\leq C_{x,y}\left(|x_t - x_0| + |y_t - y_0|\right) + 
\end{align*}

\section{Order of the approximation $\bar{x}_t$}

@@

>>>>

In order to have a Markovian representation of the HJM, we will assume that the volatility is separable, i.e.
\begin{equation}
\sigma(t,T)= h(t)g(T).
\end{equation}
with $g$ a positive time-dependent function and $h$ a non-negative process. In addition
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Throught this paper, we assume the following hypotheses on $\eta(t,x,y)$.
\begin{hypothesis}\label{boundedness_volatility} 
The process $\eta_t$ is global Lipschitz and differentiable a.s. In addition, we will suppose that
\begin{align*}
\alpha_1 \leq \eta(t,x,y) \leq \alpha_2 \quad \forall (t,x,y) \in \mathbb{R}^{+} \times \mathbb{R} \times \mathbb{R}^{+}. \\
|\eta(t,x_2,y_2) - \eta(t,x_1,y_1)| \leq C_{x,y} \lVert (x_2-x_1,y_2-y_1)\rVert \quad \forall (t,x,y) \in \mathbb{R}^{+} \times \mathbb{R} \times \mathbb{R}^{+} 
\end{align*}
with $\lVert \cdot \rVert$ euclidean norm in $\mathbb{R}^{2}$.
\end{hypothesis}
\begin{hypothesis}\label{boundedness_reversion} 
The mean reversion function $k(\cdot)$ is positive a.s.
\end{hypothesis}
\begin{remark}
Notice that the above hypotheses have been chosen for the sake of simplicity, but they can be replaced by adequate integrability conditions.
\end{remark}
We must note, that under the hypothesis (\ref{boundedness_volatility}) $\partial_x \eta(t,x,y)$ and $\partial_y \eta(t,x,y)$ are bounded. This version of the HJM is also known as the Cheyette model, \cite{Cheyette}. It is easy to show (see \cite{Andreasen01} that
\begin{equation}
r_{ois}(t)=f_{ois}(t,t)= f_{ois}(0,t) + x_t,
\end{equation}  
where
\begin{align}\label{short_rate_cheyette}
dx_t &= (-k_t x_t + y_t)dt + \eta(t,x_t,y_t) dW_t^{\mathbb{Q}} \nonumber \\
dy_t &= (\eta^{2}_t - 2 k_t y_t) dt ,
\end{align} 
with 
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Now if we do a little bit of algebra and we use Itô formula we can show the next representation formula for $P_{ois}(t,T)$
\begin{equation}\label{bond_ois}
P_{ois}(t,T) = \frac{P_{ois}(0,T)}{P_{ois}(0,t)} \exp\left(-G(t,T)x_t - \frac{1}{2} G^{2}(t,T)y_t \right)
\end{equation}
where $G(t,T) = \int_{t}^{T} \exp\left(-\int_{t}^{u} k_s ds \right) du$. We must observe, that under the representation (\ref{estimation_forward_rate_curve}) we have that
\begin{equation}\label{bond_forward}
P_{E}(t,T)=H(t,T)P_{ois}(t,T)
\end{equation}
where $H(t,T)=\exp\left(-\int_{t}^{T}s(t,u) du \right)$.

@@

@@

\section{Order of the approximation $\bar{x}_t$}
From (\ref{short_rate_cheyette}) is easy to show that
\begin{equation*}
x_{t_a} = \int_{0}^{t_a} \exp\left(-\int_{u}^{t_a}k_u du^{\prime}\right) G(u,t_a) \eta^{2}(u,x_u,y_u) du + \int_{0}^{t_a}  \exp\left(-\int_{u}^{t_a}k_u^{\prime} du \right) \eta(u,x_u,y_u) dW_u^{\mathbb{Q}} 
\end{equation*}
Therefore, we have that
\begin{align*}
(x_{t_a} - \bar{x}_{t_a})^2 &\leq 2 \left(\int_{0}^{t_a} \exp\left(-\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right) G(u,t_a) (\eta^{2}(u,x_u,y_u) - \eta^{2}(u,x_0,y_0)) du \right)^{2} \\ 
&+ 2 \left(\int_{0}^{t_a}  \exp\left(-\int_{s}^{t_a}k_u^{\prime} du^{\prime} \right) (\eta(u,x_u,y_u)-\eta(u,\bar{x}_u,\bar{y}_u)) dW_u^{\mathbb{Q}}\right)^{2}
\end{align*}
Now, we will bound each integral term. We will call to the first term $(A)$ and the second $(B)$. From the hypothesis (\ref{boundedness_volatility}) we have that 
\begin{align*}
(A) &\leq 8 t_a \alpha^{2}_2 C^{2}_{x,y} \int_{0}^{t_a} \exp\left(-2\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right) G^{2}(u,t_a) \|(x_u -x_0,y_u-y_0)\|^{2} du \\
&\leq 16 t_a \alpha^{2}_2 C^{2}_{x,y} \int_{0}^{t_a} \exp\left(-2\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right) G^{2}(u,t_a) \|(x_u - \bar{x}_u,y_u-\bar{y}_u)\|^{2} du \\
&+ 16 t_a \alpha^{2}_2 C^{2}_{x,y} \int_{0}^{t_a} \exp\left(-2\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right) G^{2}(u,t_a) \|(\bar{x}_u - x_0,\bar{y}_u-y_0)\|^{2} du
\end{align*}
and we take $\mathbb{E}^{\mathbb{Q}}\left(\cdot\right)$ on above inequality
\begin{align*}
\mathbb{E}^{\mathbb{Q}}\left((A)\right) &\leq 16  t_a \alpha^{2}_2 C^{2}_{x,y} \int_{0}^{t_a} \exp\left(-2\int_{s}^{t_a}k_u^{\prime} du^{\prime}\right) G^{2}(u,t_a) \mathbb{E}^{\mathbb{Q}}\left(\|(\bar{x}_u -x_0,\bar{y}_u-y_0)\|^{2}\right) du \\
&+\leq 16  t_a \alpha^{2}_2 C^{2}_{x,y} \int_{0}^{t_a} \exp\left(-2\int_{s}^{t_a}k_u^{\prime} du^{\prime}\right) G^{2}(u,t_a) \mathbb{E}^{\mathbb{Q}}\left(\|(x_u - \bar{x}_u,y_u-\bar{y}_u)\|^{2}\right) du
\end{align*}
Again, from the hypotheses (\ref{boundedness_volatility}) and (\ref{boundedness_reversion}) we can show that 
\begin{equation*}
\mathbb{E}^{\mathbb{Q}}\left(\|(\bar{x}_u -x_0,\bar{y}_u-y_0)\|^{2}\right) = K h(u)
\end{equation*}
where $K$ is a constant with no dependency in $u$ and $h(u)\sim \mathcal{O}(u)$. In order to bound $(B)$, we will use the Doob's inequality and (\ref{boundedness_volatility}). Then
\begin{align*}
\mathbb{E}^{\mathbb{Q}}&\left((B)\right) \leq 2K \int_{0}^{t_a}  \exp\left(-2\int_{u}^{t_a}k_u^{\prime} du^{\prime} \right) (\eta(u,x_u,y_u)-\eta(u,\bar{x}_u,\bar{y}_u))^{2} du \\
& \leq 2K C^{2}_{x,y} \int_{0}^{t_a}  \exp\left(-2\int_{u}^{t_a}k_u^{\prime} du^{\prime} \right) \mathbb{E}^{\mathbb{Q}}\left(\|(x_u - \bar{x}_u, y_u - \bar{y}_u) \|^{2}\right) du
\end{align*}
On other hand, we have that
\begin{equation*}
\mathbb{E}^{\mathbb{Q}}\left((y_{t_a} - \bar{y}_{t_a})^{2} \right) \leq 4 \alpha^{2}_{2} C^{2}_{x,y} \int_{0}^{t_a} \exp\left(-2\int_{u}^{t_a}k_u^{\prime} du^{\prime}\right)\mathbb{E}^{\mathbb{Q}}\left(\|(x_u-\bar{x}_u,y_u-\bar{y}_u)\|^{2}\right) du    
\end{equation*}
Now, if we put all of the above together, we get that
\begin{equation*}
\mathbb{E}^{\mathbb{Q}}\left(\|(x_{t_a} - \bar{x}_{t_a},y_{t_a} - \bar{y}_{t_a})\|^{2}\right) \leq C_1H(t_a) + C_2\int_{0}^{t_a} \mathbb{E}^{\mathbb{Q}}\left(\|(x_{u} - \bar{x}_{u},y_{u} - \bar{y}_{u})\|^{2}\right) du
\end{equation*}
where $H(t_a) \mathcal{O}(t_a)$. To end, if we apply the Gronwall's lemma we have that
\begin{equation*}
\mathbb{E}^{\mathbb{Q}}\left(\|(x_{t_a} - \bar{x}_{t_a},y_{t_a} - \bar{y}_{t_a})\|^{2}\right) \leq C_1H(t_a) \exp\left(C_2 t_a\right)
\end{equation*}
